Trending assessment for nowcasts, measurements, and forecasts

### 2024-03-06 sections/03-trending.tex
Add draft for probabilistic forecasting

\begin{table}
    \centering
    \begin{tabularx}{0.75\textwidth}{l X}
        \toprule
        Application & Predicted change computation \\
        \midrule
        Measurement & $(x_{t + l} - x_t)_{t=0}^T \label{eq:diffxl_measure} $\\
        Nowcasting & $
\diffxl =
\begin{cases}
(x_{t+l|t+l} -x_{t|t+l})^T_{t=0} & \text{if } y_{t} \text{ is not known at time } t, \\
(x_{t+l|t+l} -y_{t})^T_{t=0} & \text{else}.
\end{cases} \label{eq:diffxl_nowcasting}
$ \\
        Forecasting & $
\diffxl =
\begin{cases}
(x_{t+l|t} - x_{t|t})^T_{t=0} & \text{if } y_{t} \text{ is not known at time } t, \\
(x_{t+l|t} - y_{t})^T_{t=0}  & \text{else}.
\end{cases} \label{eq:diffxl_forecasting}
$\\
        \bottomrule
    \end{tabularx}
    \caption{Computation of the predicted change in the different applications. Signals $x_{t | \tau}$ refer to values issued at $\tau$ with a target time $t$. Note the different relations of issue and target time for nowcasts and forecasts. }
    \label{tab:notation}
\end{table}

@@

\subsection{Probabilistic Evaluation}

\todo{Schreiben}

>>>>

\begin{table}
    \centering
    \begin{tabularx}{0.75\textwidth}{l X}
        \toprule
        Application & Predicted change computation \\
        \midrule
        Measurement & $(x_{t + l} - x_t)_{t=1}^T \label{eq:diffxl_measure} $\\
        Nowcasting & $
\diffxl =
\begin{cases}
(x_{t+l|t+l} -x_{t|t+l})^T_{t=1} & \text{if } y_{t} \text{ is not known at time } t, \\
(x_{t+l|t+l} -y_{t})^T_{t=1} & \text{else}.
\end{cases} \label{eq:diffxl_nowcasting}
$ \\
        Forecasting & $
\diffxl =
\begin{cases}
(x_{t+l|t} - x_{t|t})^T_{t=1} & \text{if } y_{t} \text{ is not known at time } t, \\
(x_{t+l|t} - y_{t})^T_{t=1}  & \text{else}.
\end{cases} \label{eq:diffxl_forecasting}
$\\
        \bottomrule
    \end{tabularx}
    \caption{Computation of the predicted change in the different applications. Signals $x_{t | \tau}$ refer to values issued at $\tau$ with a target time $t$. Note the different relations of issue and target time for nowcasts and forecasts. }
    \label{tab:notation}
\end{table}

@@

\newpage
\subsection{Probabilistic Evaluation}\label{subsec:probabilistic}

In nowcasting and forecasting, there is a shift from point predictions to probabilistic predictions (see, for example, Sections~\ref{sec:application-covid} and~\ref{sec:application-eda} or the review in~\cite{Gneiting2014}).
Probabilistic predictions include a point estimate and information on forecast uncertainty and quantiles simultaneously.
In this section, we extend the trending assessment to probabilistic predictions.

Probabilistic predictions can have the form of a \ac{cdf}, \ac{pdf}, or quantiles.
The \ac{cdf} is the most general and can be used to derive the others, given they exist.
Let us first assume that the prediction is a \ac{cdf} and the lag is large enough such that the true value is known when issuing the forecast or nowcast.
See the end of the section for predictions given as quantiles or for unknown true values.

Let $F_{t | \tau} (x)$ denote the predicted \ac{cdf} of forecast time $t$ and issue time $\tau$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
For the trending assessment, we contrast the probability of positive change to the occurrence of positive changes. 
The probability $p_t$ of a positive change is given by one minus the \ac{cdf} of the prediction at the true value. \unsure{Andere Formulierung für: one minus?}
As for the differences in Section~\ref{subsec:notation}, the computation differs slightly for nowcasts and forecasts.
For nowcasts, the probability is calculated by
\begin{equation*}
    p_t = 1 - F_{t+l | t+l} (y_{t})\quad t = 1, \dots, T-l;
\end{equation*}
while for forecasts, it is given by
\begin{equation*}
    p_t = 1 - F_{t+l | t} (y_{t})\quad t = 1, \dots, T-l.
\end{equation*}
Let $z_t$ denote the indicator that the occurred change at time $t$ is positive, that is,
\begin{equation*}
    z_t = \ind{\diffyt > 0} \quad t = 1, \dots, T-l.
\end{equation*}
The probabilistic trending evaluation is then based on the probabilities $\mathbf{p} = (p_t)_{t=1}^{T-l}$ and the realized changes $\mathbf{z} = (z_t)_{t=1}^{T-l}$.
The predictive power of $\mathbf{p}$ for $\mathbf{z}$ can be assessed using the methods of probabilistic dichotomous forecast evaluation.
Dichotomous forecasts predict a binary outcome, for example, a positive or negative change.
Numerical measures for forecast quality are scoring rules.

The \ac{bs} is a widely used scoring rule for dichotomous, probabilistic forecasts~\parencite{Brier1950}.
It is defined as
\begin{equation*}
    BS (\mathbf{p}, \mathbf{z}) = \frac{1}{T-l} \sum_{t=1}^{T-l} (p_t - z_t)^2.
\end{equation*}
The \ac{bs} assesses the calibration and sharpness of the forecast and observation simultaneously~\parencite{Ranjan2010,Mitchell2011}.
Calibration refers to the statistical consistency of forecasts and observations, while sharpness measures the spread of the forecast distribution.
A smaller spread and thus higher sharpness is preferable, as it indicates a greater confidence in the prediction.
The basic paradigm of probabilistic forecasting is to maximize the sharpness, subject to being calibrated~\parencite{Gneiting2014}.
These two conflicting aspects cannot be distinguished in the scoring rule.

Graphical models are a standard tool for assessing solely the calibration of probabilistic forecasts.
In dichotomous forecasting, the reliability diagram is frequently used~\parencite{Ranjan2010}.
The reliability diagram plots the observed frequency of the positive outcome against the forecasted probability.
Ideally, the forecasted probability equals the observed frequency, and the reliability diagram is a 45-degree line.
Local deviations from the 45-degree line indicate miscalibration of the forecast for specific probabilities.
Thus, the reliability visualizes local and overall calibration simultaneously.
For an example of a reliability diagram, see Section~\ref{sec:application-eda}.

If the forecasts or nowcasts are given as quantiles, $p_t$ can be computed by interpolations between the quantiles.
Let $q_p$ denote the quantiles for forecast time $t+l$ for even-spaced probabilities $p \in \{1/P, \dots, (p-1) / P\}$ ($P \in \N \setminus \{1, 2\}$) and $y_t$ the true value at time $t$.
Then, the probability $\bar{p}_t$ of a negative change is between $p^{\star}$ and $p^{\star} + 1/P$ for
\begin{equation*}
    p^{\star} = \max \{p \in \{1/P, \dots, (p-1) / P\}: q_p \leq y_t\} , \text{if} q_{1/p} \leq y_t \leq q_{1 - 1/p}.
\end{equation*}
The quantiles do not determine the location within the interval.
Under the assumption of a uniform distribution within the quantile interval, the probability of a negative change is
\begin{equation*}
    \bar{p}_t = \frac{y_t - q_{p^\star}}{P (q_{p^{\star} + 1} - q_{p^{\star}})} + p^{\star}.
\end{equation*}
If $y_t$ is smaller than the smallest quantile $q_{1/p}$ or larger than the largest quantile, we assume the probability mass to be uniformly distributed on an interval of the same length as the nearest interval specified by the quantiles.
This yields
\begin{equation*}
\bar{p}_t = \begin{cases}
    \max \{\frac{1}{P} - \frac{q_{p^\star} - y_t}{P (q_{p2/P} - q_{1/P})}, 0\} &, \text{if } y_t < q_{1/p}, \\
    \min \{\frac{1}{P} - \frac{y_t - q_{(P-1)/P}}{P (q_{(P-1)/P} - q_{(P-2)/P})}, 1\} &, \text{if } y_t > q_{1 - 1/p}, \\
    \frac{y_t - q_{p^\star}}{P (q_{p^{\star} + 1} - q_{p^{\star}})} + p^{\star} &, \text{otherwise.}
\end{cases}
\end{equation*}
The probability of positive change is $p_t = 1 - \bar{p}_t$.

If the true value is unknown at lag $l$, the probabilities $p_t$ can be computed by integration.
Let for two nowcasts the distributions be given by \acp{pdf} $f_{t+l|t+l}$ and $f_{t|t+l}$ with \acp{cdf}  $F_{t+l|t+l}$ and $F_{t|t+l}$.
Then, the probability of a negative change can be computed by
\begin{align*}
    \bar{p}_t
        &= \int_{\begin{subarray}{l}x_1, x_2 \in \R: \\ x_2 < x_1\end{subarray}} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \\
        &= \int_{x_1 \in \R} \int_{-\infty}^{x_1} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \\
        &= \int_{x_1 \in \R} f_{t|t+l} (x_1) F_{t+l|t+l} (x_1)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1.
\end{align*}
Thereby, the distributions are assumed to be independent.
As a Monte Carlo approximation, the probability can also be computed by sampling from $f_{t+l|t+l}$ and $f_{t|t+l}$ and calculating the fraction of negative changes.
For forecasts, the indexes have to be shifted.
If no \acp{pdf} are available, they can be estimated from the \ac{cdf} or quantiles, or samples can be generated based on the \ac{cdf} or quantiles.

### 2024-03-06 sections/04-application.tex
Add draft for probabilistic forecasting

@@

@@

Trending ability is consistent for the two lags, with the models' relative trending ability evolving similarly for the two lags.
The models' trending ability is generally higher for the smaller lag, but the differences are minor, and confidence intervals overlap.

@@

@@

\begin{table}
    \centering
    \caption{Data summary of the \ac{abp} and \ac{nbp} measurements for the lags one minute, five minutes, and 15 minutes.}
    \label{tab:app-mimic-data-summary}
\end{table}

>>>>

Figure~\ref{fig:app-eda-prob} visualizes the probabilistic trending evaluation for the same subset of models.
The \acf{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the lags 3 and 7 days are shown Figures~\ref{fig:app-eda-prob-rel-3} and \ref{fig:app-eda-prob-rel-7}.
The \acp{bs} are smallest and close for NBI-2 and Poisson-2 for both lags, while the Brier scores for the other models are larger and differ more.
The qreg-1 model has the highest \ac{bs} for both lags.
The reliability diagrams of GBM-2 and NBI-2 are also close and show overall a too small fraction of increases for the predicted probability.
For the other models, the reliability diagrams show rather a too large fraction of increases for the corresponding predicted probability.

@@

\begin{figure}
    \begin{subfigure}{0.32\textwidth}
    \tiny
    \input{plots/ed_arrival/70_brier_scores}
    \caption{Brier Scores for the different models and lags.}\label{fig:app-eda-prob-brier}
    \end{subfigure}\hspace{0.01\textwidth}%
    \begin{subfigure}[t]{0.32\textwidth}
    \includegraphics{plots/ed_arrival/60_reliability_diagram_lag_3}
    \caption{Reliability diagram for lag 3 days.}\label{fig:app-eda-prob-rel-3}
    \end{subfigure}\hspace{0.01\textwidth}%
    \begin{subfigure}[t]{0.32\textwidth}
    \includegraphics{plots/ed_arrival/60_reliability_diagram_lag_7}
    \caption{Reliability diagram for lag 7 days.}\label{fig:app-eda-prob-rel-7}
    \end{subfigure}
    \caption{Probabilistic trending evaluation for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the lags 3 and 7 days.}
    \label{fig:app-eda-prob}
\end{figure}

@@

The trending ability is consistent for the two lags, with the models' relative trending ability evolving similarly for the two lags.
The models' trending ability is generally higher for the smaller lag, but the differences are minor, and confidence intervals overlap.

@@

The results of probabilistic trending evaluation \todo{Fertig schreiben}

@@

\begin{table}
    \centering
    \missingfigure{MIMIC Data summary}
%    \input{plots/mimic/}
    \caption{Data summary of the \ac{abp} and \ac{nbp} measurements for the lags one minute, five minutes, and 15 minutes.}
    \label{tab:app-mimic-data-summary}
\end{table}
